---
title: "LTERWG SOM time series"
authors: "Derek Pierson & Jessica Moore"
date: "November 14, 2018"
output:
  rmarkdown::html_document:
    theme: lumen
---

## Load Libraries and SOMWG compiled data (aka tarball)
```{r warning = FALSE, message = FALSE}
library(dplyr)
library(ggplot2)

#??? Could we bring in data from a remote source (= common address)
data.all <- read.csv("C:/Google Drive/Code/R/SOM_tarball_11-9-2018.csv", stringsAsFactors = FALSE)
#data.all <- readRDS("/Users/JM 1/Downloads/somCompositeData_2018-11-09T17_34_23.rds") 

data.all$full_name <- paste0(data.all$site_code," ",data.all$location_name)  #Concatenate strings
data.all <- data.all %>% mutate(full_name = gsub("NA", "", full_name))  #Remove NA from strings
```

## Fix and combine YEAR data from columns 'observation_date' and 'observation_date-1'
```{r warning = FALSE}
#get first four characters from observation_date
data.all$year<-substr(data.all$observation_date, start=1, stop=4)
data.all$year<-as.numeric(data.all$year)

#get first four characters from observation_date_1
data.all$year_1<-substr(data.all$observation_date_1, start=1, stop=4)
data.all$year_1<-as.numeric(data.all$year_1)

#merge year and year_1 using ifelse
data.all$year_all<-ifelse(data.all$year=="NA",data.all$year_1,data.all$year)

```

## Pull YYYY values from treatment level columns ('L1' 'L2' 'L3' etc.)
```{r warning = FALSE}
#getting obs year out of level 1
data.all$yr_L1 <- as.numeric(data.all$L1)
data.all <- data.all %>% mutate(yr_L1 = "is.na<-"(yr_L1, yr_L1 < 1900)) %>% mutate(yr_L1 = "is.na<-"(yr_L1, yr_L1 > 2020)) 
data.all$year_all<-ifelse(data.all$year_all=="NA",data.all$yr_L1,data.all$year_all)

#getting obs year out of level 2
data.all$yr_L2 <- as.numeric(data.all$L2)
data.all <- data.all %>% mutate(yr_L2 = "is.na<-"(yr_L2, yr_L2 < 1900)) %>% mutate(yr_L2 = "is.na<-"(yr_L2, yr_L2 > 2020)) 
data.all$year_all<-ifelse(data.all$year_all=="NA",data.all$yr_L2,data.all$year_all)

#getting obs year out of level 3
data.all$yr_L3 <- as.numeric(data.all$L3)
data.all <- data.all %>% mutate(yr_L3 = "is.na<-"(yr_L3, yr_L3 < 1900)) %>% mutate(yr_L3 = "is.na<-"(yr_L3, yr_L3 > 2020)) 
data.all$year_all<-ifelse(data.all$year_all=="NA",data.all$yr_L3,data.all$year_all)

#getting obs year out of level 4
data.all$yr_L4 <- as.numeric(data.all$L4)
data.all <- data.all %>% mutate(yr_L4 = "is.na<-"(yr_L4, yr_L4 < 1900)) %>% mutate(yr_L4 = "is.na<-"(yr_L4, yr_L4 > 2020)) 
data.all$year_all<-ifelse(data.all$year_all=="NA",data.all$yr_L4,data.all$year_all)
```

## Clean up and check data in compiled 'year_all' column
```{r warning = FALSE}
#remove 3 digit values in year_all
data.all <- data.all %>% mutate(year_all = "is.na<-"(year_all, year_all < 1900))

#check for unique 4 digit values in year
unique(data.all$year_all)
```


# Histogram of SOMWG data by year
```{r warning = FALSE}
ggplot(data.all, aes(x=year_all))+
  geom_histogram(aes(fill=site_code))+
  labs(x="Observation Year", y="Count of All Data")+
  facet_grid(.~network)
```


# Data exploration

##SOC Stocks

### Clean up soc data values
```{r warning = FALSE}
#turn negative soc stocks into NA, new dataframe called data.soc
data.soc<- data.all %>% filter(!is.na(soc_stock)) %>% filter(soc_stock>0)
```

### SOC stock data exploration
```{r warning = FALSE}
ggplot(data.soc, aes(x=year_all, y=soc_stock))+
  geom_point(aes(color=site_code, shape=network))+
  scale_y_continuous(limits=c(0,30000))+
  labs(x="Observation Year", y="SOC stock")
```

## SOC Percents

### Clean up SOC percent values
```{r warning = FALSE}
#filter out negative soc%
data.soc.percent<-data.all %>% filter(!is.na(soc)) %>% filter(soc>0) %>% filter(soc<1)
```

### SOC percent data exploration
```{r warning = FALSE}
soc.prop<-ggplot(data.soc.percent, aes(x=year_all, y=soc))+
  geom_point(aes(color=site_code, shape=network))+
  #scale_y_continuous(limits=c(0,30000))+
  labs(x="Observation Year", y="SOC proportion")
soc.prop
```

```{r warning = FALSE}
soc.per<-ggplot(data.soc.percent, aes(x=year_all, y=soc))+
  geom_point(aes(color=site_code, shape=network))+
  #scale_y_continuous(limits=c(0,30000))+
  labs(x="Observation Year", y="SOC %")
```

```{r warning = FALSE}
soc.per<-ggplot(data.soc.percent, aes(x=year, y=soc))+
  geom_point(aes(color=site_code))+
  facet_grid(.~network)
soc.per
```

# Site-specific time series analyses

##UMBS
```{r warning = FALSE}
#Isolate UMBS DIRT data as new dataframe, filter for CONTROL only
data.soc.percent.umbs.c<-data.soc.percent %>% filter(site_code=="UMBS") %>% filter(tx_L1=="C")

#Plot SOC by YEAR
soc.per<-ggplot(data.soc.percent.umbs.c, aes(x=year_all, y=soc))+
  geom_point()+
  #scale_y_continuous(limits=c(0,30000))+t
  labs(x="Observation Year", y="SOC in Control Plots")
soc.per
```



### UMBS temp, not year
soc.per<-ggplot(data.soc.percent.umbs.c, aes(x=, y=soc))+
  geom_point()+
  #scale_y_continuous(limits=c(0,30000))+t
  labs(x="Observation Year", y="SOC in Control Plots")
soc.per